---
title: "Deep Learning Algorithms"
output: html_document
---


# Packages:
```{r}

suppressMessages(library(ggplot2))
suppressMessages(library(caret))
suppressMessages(library(randomForest))
suppressMessages(library(ggpubr))
suppressMessages(library(MuMIn))
suppressMessages(library(dplyr))
suppressMessages(library(viridis))
suppressMessages(library(grid))
suppressMessages(library(car))
suppressMessages(library(respR))
suppressMessages(library(rfviz))
suppressMessages(library(tidyverse))
suppressMessages(library(olsrr))


```


# Data upload function:
```{r}

# Upload the data from a total data file:
# total_data <- read.csv("Total_Sat_Data_All_Var.csv")

# Create a function for reading and preprocessing data from the file:
Read_fn <- function(file) {
  
  
  data <- read.csv(file)  # Read the csv
  
  data <- data %>% 
    mutate(HBE = if_else(Chl > 2.5, 1, 0))
  
  
  # Total dataset:
  all_data <- data
  
  # Only Pacific data:
  Pacific_data = data[data$Area=='ESNP', ]
  #Pacific_data = Pacific_data[Pacific_data$Test_Cruise!='Summer_2022', ]
  Pacific_data = subset(Pacific_data, select=-c(Area))  # Remove redundant area variable
  Pacific_data_test = Pacific_data
  #Pacific_data <- na.omit(Pacific_data)
  
  # Only Arctic data:
  Arctic_data = data[data$Area=='BS-CB', ]
  Arctic_data = subset(Arctic_data, select=-c(Area))
  #Arctic_data <- na.omit(Arctic_data)
  
  
  # Only Bering Strait data:
  Bering_data = data[data$Area=='BE-CH', ]
  Bering_data = subset(Bering_data, select=-c(Area))
  #Bering_data <- na.omit(Bering_data)
  
  # Merge into a list:
  dataset <- list(all_data, Pacific_data, Arctic_data, Bering_data, Pacific_data_test)

  
  return(dataset)
  
}


# Input required files into function (from previous file):
Pacific_data <- as.data.frame(Read_fn("MODEL_DATA_GLOBAL_FINAL_Woutlier.csv")[2])
Pacific_data_test <- as.data.frame(Read_fn("MODEL_DATA_GLOBAL_FINAL_Woutlier.csv")[5])
Arctic_data <- as.data.frame(Read_fn("MODEL_DATA_GLOBAL_FINAL_Woutlier.csv")[3])
Bering_data <- as.data.frame(Read_fn("MODEL_DATA_GLOBAL_FINAL_Woutlier.csv")[4])
all_data <- as.data.frame(Read_fn("MODEL_DATA_GLOBAL_FINAL_Woutlier.csv")[1])


```


# Create variables for Apparent Oxygen Utilization (AOU) and Solar Radiaiton Dose (SRD):
```{r}

# SRD:

Pacific_data$Sat_PAR_watts <- Pacific_data$Sat_PAR * 0.21739130434  # Convert PAR to Wm2
all_data$Sat_PAR_watts <- all_data$Sat_PAR * 0.21739130434  # Convert PAR to Wm2

Pacific_data$SRD <- ((Pacific_data$Sat_PAR_watts / (Pacific_data$AC * Pacific_data$MLD)) * (1 - exp((Pacific_data$AC*-1) * Pacific_data$MLD)))  # Calculate SRD

all_data$SRD <- ((all_data$Sat_PAR_watts / (all_data$AC * all_data$MLD)) * (1 - exp((all_data$AC*-1) * all_data$MLD)))  # Calculate SRD

# AOU
Pacific_data$DO_eq <- convert_DO(Pacific_data$O2_equil_conc, # data to convert
                   from = "umol/Kg",     # oxygen unit to convert from
                   to = "mg/L",    # oxygen unit to convert to
                   t = Pacific_data$Temperature,            # in C
                   S = Pacific_data$Salinity            # in ppt
                   )         # in bar

Pacific_data$DO_meas <- convert_DO(Pacific_data$DO, # data to convert
                   from = "ml/L",     # oxygen unit to convert from
                   to = "mg/L",    # oxygen unit to convert to
                   t = Pacific_data$Temperature,            # in C
                   S = Pacific_data$Salinity            # in ppt
                   )         # in bar


# Remove nans from DO column from total dataframe:
all_data <- all_data[!is.na(all_data$O2_equil_conc),]


all_data$DO_eq <- convert_DO(all_data$O2_equil_conc, # data to convert
                   from = "umol/Kg",     # oxygen unit to convert from
                   to = "mg/L",    # oxygen unit to convert to
                   t = all_data$Temperature,            # in C
                   S = all_data$Salinity            # in ppt
                   )         # in bar

all_data$DO_meas <- convert_DO(all_data$DO, # data to convert
                   from = "ml/L",     # oxygen unit to convert from
                   to = "mg/L",    # oxygen unit to convert to
                   t = all_data$Temperature,            # in C
                   S = all_data$Salinity            # in ppt
                   )         # in bar

# Calculate produced DO
Pacific_data$DO_Prod <- as.numeric(Pacific_data$DO_meas - Pacific_data$DO_eq)
all_data$DO_Prod <- as.numeric(all_data$DO_meas - all_data$DO_eq)



```

# Sort and index data with relevant variables and all numeric:
```{r}

pacific.learning.ds <- data.frame(Pacific_data %>% # all vars of interest
  dplyr::select(TEP, Log_Chl, POC, DO_Prod, #DO,
                Temperature, MLD, WS, Salinity,
                Nitrate, Phosphate, Silicate, SRD, Sat_PAR, Seasons
                ))

all.learning.ds <- data.frame(all_data %>% # all vars of interest
  dplyr::select(TEP, Log_Chl, POC, DO_Prod, #DO,
                Temperature, MLD, WS, Salinity,
                Nitrate, Phosphate, Silicate, Seasons, Area
                ))

all.learning.ds$Area <- as.numeric(as.factor((all.learning.ds$Area))) # Transform categorical variable to numeric

```

# Random Forest call:
```{r}

# Told RF models:
rf <- randomForest(TEP~., data=pacific.learning.ds, na.action = na.omit, importance = TRUE)  # Pacific
rf.all <- randomForest(TEP~., data=all.learning.ds, na.action = na.omit, importance = TRUE)  # All data


# Create a plot to visualize variable importance in the model:
imp <- varImpPlot(rf, type = 1)
imp.all <- varImpPlot(rf.all, type = 1)


# this part just creates the data.frame for the plot part
imp <- as.data.frame(imp)
imp$varnames <- rownames(imp) # row names to column
imp$varnames <- c("Chlorophyll a", "POC", "AOU", "Temperature", "MLD", "Wind speed", "Salinity", 
                  "Nitrate", "Phosphate", "Silicic Acid", "SRD", "24h-PAR", "Season")
rownames(imp) <- NULL  
imp$IncMSE.percent <- imp$`%IncMSE`
imp$driver <- c("Source", "Source", "Source", "Direct Environ.",
                "Indirect Environ.", "Indirect Environ.", "Indirect Environ.", 
                "Direct Environ.", "Direct Environ.", "Direct Environ.", # nuts
                "Indirect Environ.", "Indirect Environ.", "Indirect Environ.")  #  driver categories
 
RFplot <- ggplot(imp, aes(x=reorder(varnames, IncMSE.percent), 
                          y = IncMSE.percent, fill = driver, colour = driver)) + 
  geom_point(size = 5, pch = 21) +
  geom_segment(aes(x = varnames, xend = varnames, y = 0, yend = IncMSE.percent)) +
  scale_color_discrete(name="Variable Group") +
  ylab("") +
  xlab("") +
  scale_fill_manual(values = c("blue", "darkcyan", "yellow4")) +
  scale_colour_manual(values = c("blue", "darkcyan", "yellow4")) +
  scale_x_discrete(limits = c(
                              "Salinity", "SRD", "Wind speed", "24h-PAR", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
  ylim(0, 25) +
  coord_flip() +
  theme_bw() +
  ggtitle("ESNP: Random Forest Regression") +
  theme(panel.grid.major.y = element_blank(), legend.title = element_blank(),
        plot.title = element_text(size = 6, face = "bold"))


# All data
imp.all <- as.data.frame(imp.all)
imp.all$varnames <- rownames(imp.all) # row names to column
imp.all$varnames <- c("Chlorophyll a", "POC", "AOU", "Temperature", "MLD", "Wind speed", "Salinity", 
                  "Nitrate", "Phosphate", "Silicic Acid", "Season", "Area")
rownames(imp.all) <- NULL  
imp.all$IncMSE.percent <- imp.all$`%IncMSE`
imp.all$driver <- c("Source", "Source", "Source", "Direct Environ.",
                "Indirect Environ.", "Indirect Environ.", "Indirect Environ.", 
                "Direct Environ.", "Direct Environ.", "Direct Environ.", # nuts
                "Indirect Environ.", "Indirect Environ.")  #  driver categories
 
RFplot.all <- ggplot(imp.all, aes(x = varnames, y = IncMSE.percent, fill = driver, colour = driver)) + 
  geom_point(size = 5, pch = 21) +
  geom_segment(aes(x = varnames, xend = varnames, y = 0, yend = IncMSE.percent)) +
  scale_color_discrete(name="Variable Group") +
  ylab("Mean Decrease Accuracy (%)") +
  xlab("") +
  scale_fill_manual(values = c("blue", "darkcyan", "yellow4")) +
  scale_colour_manual(values = c("blue", "darkcyan", "yellow4")) +
  scale_x_discrete(limits = c(
                              "Area", "Salinity", "Wind speed", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
  ylim(0, 25) +
  coord_flip() +
  theme_bw() +
  ggtitle("Total Scope: Random Forest Regression") +
  theme(panel.grid.major.y = element_blank(), legend.title = element_blank(), 
        legend.position = "none", plot.title = element_text(size = 6, face = "bold"))
  
  

ggarrange(RFplot, RFplot.all, common.legend = TRUE)

```

# Visualize the RF:
```{r}

# data_x <- data.frame(pacific.learning.ds %>% # all vars of interest
#   dplyr::select(Log_Chl, POC, DO_Prod, #DO,
#                 Temperature, MLD, WS, Salinity,
#                 Nitrate, Phosphate, Silicate, SRD, Sat_PAR, Seasons
#                 ))
# 
# data_y <- pacific.learning.ds$TEP
# 
# rfprep <- rf_prep(data_x, data_y)
# varImpPlot(rfprep$rf)
# 
# bcrf <- rf_viz(rfprep, input=TRUE, imp=TRUE, cmd=TRUE)


```




# Create a random seed, training and testing sets:
```{r}

set.seed(222)
ind <- sample(2, nrow(pacific.learning.ds), replace = TRUE, prob = c(0.8, 0.2))
train <- pacific.learning.ds[ind==1,]
test <- pacific.learning.ds[ind==2,]

```

# Predictions and Confusion matrix:
```{r}

# Training RF model:
rf <- randomForest(TEP~., data=train, proximity=TRUE, type = "regression") 

train$tep.predictions <- predict(rf, train)
train$id <- "train"

test$tep.predictions <- predict(rf, test)
test$id <- "test"

rf.data <- rbind(train, test)

ggplot(rf.data, aes(TEP, tep.predictions, colour = id)) +
  geom_point() +
  ylim(0, 150) +
  xlim(0, 150) +
  geom_abline(slope = 1, intercepot = 0, lty = 'dashed') +
  theme_bw()

MAE(test$tep.predictions, test$TEP)
RMSE(test$tep.predictions, test$TEP)


# p2 <- as.factor(predict(rf, test))
# confusionMatrix(factor(tests[,1]), factor(tests[,2]))
# 
# 


```

# Predict historical changes with the RF model:
```{r}

# # Upload Line P data:
# spring_hist <- read.csv("Spring_hist_withO2.csv")
# summer_hist <- read.csv("Summer_hist_withO2.csv")
# winter_hist <- read.csv("Winter_hist_withO2.csv")



```

#  Multiple linear regression models: ESNP
```{r}

# First, derive multiple linear regression models (MLRs) from the dataset:

# Final model versions:
ESNP.final.model <- lm(TEP ~ Log_Chl + Temperature + Nitrate + DO_Prod + MLD, 
                          data = as.data.frame(scale(
                            pacific.learning.ds[, sapply(pacific.learning.ds, is.numeric)]))) #ESNP
ESNP.alt.model <- lm(TEP ~ Log_Chl + Temperature + DO_Prod + MLD #+ Nitrate 
                     + POC + WS + Phosphate, 
                          data = as.data.frame(scale(
                            pacific.learning.ds[, sapply(pacific.learning.ds, is.numeric)]))) #ESNP

# List models:
model_list <- list(ESNP.final.model, ESNP.alt.model)
model_names <- c("ESNP", "ESNP")
model_level <- c("ESNP", "ESNP-total")


CoefficientPlot <- function(models, alpha = 0.1, 
                            modelnames = model_names, modellevel = model_level){
  # models must be a list()

  Multiplier <- qnorm(1 - alpha / 2)
  CoefficientTables <- lapply(models, function(x){summary(x)$coef})
  TableRows <- unlist(lapply(CoefficientTables, nrow))

  if(modelnames[1] == ""){
    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
    } else {
    ModelNameLabels <- rep(modelnames, TableRows)
    }
  
    if(modellevel[1] == ""){
    ModelLevelLabels <- rep(paste("Level", 1:length(TableRows)), TableRows)
    } else {
    ModelLevelLabels <- rep(modellevel, TableRows)
    }

  MatrixofModels <- cbind(do.call(rbind, CoefficientTables), ModelNameLabels, ModelLevelLabels)
  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels, ModelLevelLabels))
  colnames(MatrixofModels) <- c("IV", "Estimate", "StandardError", 
                                "TValue", "PValue", "ModelName", "ModelLevel")
  MatrixofModels$ModelLevel = factor(MatrixofModels$ModelLevel, 
                                    levels = c("ESNP", "ESNP-total"))
  MatrixofModels$ModelName = factor(MatrixofModels$ModelName, 
                                    levels = c("ESNP", "ESNP-total"))
  #MatrixofModels$IV <- factor(MatrixofModels$IV, levels = MatrixofModels$IV)
  #MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})
  MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})

  OutputPlot <- qplot(IV, Estimate, 
                      ymin = Estimate - Multiplier * StandardError,
                      ymax = Estimate + Multiplier * StandardError,
                      data = MatrixofModels, geom = "pointrange", 
                      ylab = NULL, xlab = NULL, 
                      group = ModelLevel, shape = ModelLevel, alpha = ModelLevel,
                      colour = driver, fill = driver, size = I(1)) +
    geom_hline(yintercept = 0, lwd = I(7/12), colour = I(hsv(0/12, 7/12, 7/12)), 
              alpha = I(5/12)) +
    coord_flip() + theme_bw() +
    theme(panel.grid.major.y = element_blank(), legend.title = element_blank()) +
    ggtitle("ESNP: Multiple Linear Regression") +
    #geom_point(colour = 'black') +
    scale_x_discrete(limits = c(
                                "Salinity", "SRD", "WS", "Sat_PAR", "MLD",
                                "Silicate", "Phosphate", "Nitrate", "Temperature",
                                "DO_Prod", "POC", "Log_Chl"), 
                     labels = c(
                              "Salinity", "SRD", "Wind speed", "24h-PAR", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
    scale_shape_manual(values = c("ESNP" = 19, "ESNP-total" = 1)) + 
    scale_fill_manual(values = c("Source" = "yellow4", "Direct Environ." = "blue", 
                                 "Indirect Environ." = "darkcyan")) +
    scale_colour_manual(values = c("Source" = "yellow4", "Direct Environ." = "blue", 
                                 "Indirect Environ." = "darkcyan")) +
    scale_alpha_manual(values = c(1, 1)) +
    theme(legend.title = element_blank(), legend.position = "top", 
          axis.text.y = element_blank(), plot.title = element_text(size = 6, face = "bold")) +
    ylim(-0.3, 0.75) +
    ylab("") +
    xlab("")
  
  return(OutputPlot)
}


MLRplot <- CoefficientPlot(model_list)

# Choose the color scheme corresponding to each parameter:
driver <- c("Source", "Source", "Direct Environ.", "Direct Environ.",
                "Source", "Indirect Environ.", "Indirect Environ.",
                "Source", 
                "Direct Environ.", "Source", "Indirect Environ.", # nuts
                #"Direct Environ.", 
                "Source", "Indirect Environ.",
                "Direct Environ.")  #  driver categories

# order: Log_Chl, Temperature, Nitrate, DO_Prod, MLD, Log_Chl, Temperature, DO_Prod, MLD, Nitrate, POC, WS,
# Phosphate

# print the pdf:
pdf(file = "Coefficients.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches


esnp.plots <- ggarrange(RFplot, MLRplot, common.legend = TRUE)


dev.off()


```

# Compare Random Forest model with multiple linear regression models: Different style:
```{r}


# Restyle the RF:
RFplot2 <- ggplot(imp, aes(x=reorder(varnames, IncMSE.percent), 
                          y = IncMSE.percent, fill = IncMSE.percent, colour = IncMSE.percent)) + 
  geom_point(size = 5, pch = 21) +
  geom_segment(aes(x = varnames, xend = varnames, y = 0, yend = IncMSE.percent)) +
  scale_fill_viridis(option = 'D') +
  scale_colour_viridis(option = 'D') +
  ylab("Mean Decrease Accuracy (%)") +
  xlab("") +
  scale_x_discrete(limits = c(
                              "Salinity", "SRD", "Wind speed", "24h-PAR", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
  coord_flip() +
  theme_bw() +
  ggtitle("ESNP: Random Forest Regression") +
  theme(panel.grid.major.y = element_blank(), legend.title = element_blank(), legend.position = "bottom")

##--------

# Final model versions:
ESNP.final.model <- lm(TEP ~ Log_Chl + Temperature + Nitrate + DO_Prod + MLD, 
                          data = as.data.frame(scale(
                            pacific.learning.ds[, sapply(pacific.learning.ds, is.numeric)]))) #ESNP
ESNP.alt.model <- lm(TEP ~ Log_Chl + Temperature + DO_Prod + MLD #+ Nitrate 
                     + POC + WS + Phosphate, 
                          data = as.data.frame(scale(
                            pacific.learning.ds[, sapply(pacific.learning.ds, is.numeric)]))) #ESNP

# List models:
model_list <- list(ESNP.final.model, ESNP.alt.model)
model_names <- c("ESNP", "ESNP")
model_level <- c("ESNP", "ESNP-total")


CoefficientPlot <- function(models, alpha = 0.1, 
                            modelnames = model_names, modellevel = model_level){
  # models must be a list()

  Multiplier <- qnorm(1 - alpha / 2)
  CoefficientTables <- lapply(models, function(x){summary(x)$coef})
  TableRows <- unlist(lapply(CoefficientTables, nrow))

  if(modelnames[1] == ""){
    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
    } else {
    ModelNameLabels <- rep(modelnames, TableRows)
    }
  
    if(modellevel[1] == ""){
    ModelLevelLabels <- rep(paste("Level", 1:length(TableRows)), TableRows)
    } else {
    ModelLevelLabels <- rep(modellevel, TableRows)
    }

  MatrixofModels <- cbind(do.call(rbind, CoefficientTables), ModelNameLabels, ModelLevelLabels)
  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels, ModelLevelLabels))
  colnames(MatrixofModels) <- c("IV", "Estimate", "StandardError", 
                                "TValue", "PValue", "ModelName", "ModelLevel")
  MatrixofModels$ModelLevel = factor(MatrixofModels$ModelLevel, 
                                    levels = c("ESNP", "ESNP-total"))
  MatrixofModels$ModelName = factor(MatrixofModels$ModelName, 
                                    levels = c("ESNP", "ESNP-total"))
  #MatrixofModels$IV <- factor(MatrixofModels$IV, levels = MatrixofModels$IV)
  #MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})
  MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})

  OutputPlot <- qplot(IV, Estimate, 
                      ymin = Estimate - Multiplier * StandardError,
                      ymax = Estimate + Multiplier * StandardError,
                      data = MatrixofModels, geom = "pointrange", 
                      ylab = NULL, xlab = NULL, 
                      group = ModelLevel, shape = ModelLevel, alpha = ModelLevel,
                      colour = abs(Estimate), fill = abs(Estimate), size = I(1)) +
    geom_hline(yintercept = 0, lwd = I(7/12), colour = I(hsv(0/12, 7/12, 7/12)), 
              alpha = I(5/12)) +
    coord_flip() + theme_bw() +
    theme(panel.grid.major.y = element_blank(), legend.title = element_blank()) +
    ggtitle("Multiple Linear Regression") +
    #geom_point(colour = 'black') +
    scale_x_discrete(limits = c(
                                "Salinity", "SRD", "WS", "Sat_PAR", "MLD",
                                "Silicate", "Phosphate", "Nitrate", "Temperature",
                                "DO_Prod", "POC", "Log_Chl"), 
                     labels = c(
                              "Salinity", "SRD", "Wind speed", "24h-PAR", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
    scale_shape_manual(values = c("ESNP" = 19, "ESNP-total" = 1)) + 
    scale_fill_viridis(option = 'D') +
    scale_colour_viridis(option = 'D') +
    xlab("") +
    scale_alpha_manual(values = c(1, 1)) +
    theme(legend.title = element_blank(), legend.position = "bottom")
  
  return(OutputPlot)
}


MLRplot <- CoefficientPlot(model_list)

# Choose the color scheme corresponding to each parameter:
driver <- c("Source", "Source", "Direct Environ.", "Direct Environ.",
                "Source", "Indirect Environ.", "Indirect Environ.",
                "Source", 
                "Direct Environ.", "Source", "Indirect Environ.", # nuts
                #"Direct Environ.", 
                "Source", "Indirect Environ.",
                "Direct Environ.")  #  driver categories

# order: Log_Chl, Temperature, Nitrate, DO_Prod, MLD, Log_Chl, Temperature, DO_Prod, MLD, Nitrate, POC, WS,
# Phosphate

# print the pdf:
pdf(file = "Coefficients.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 6) # The height of the plot in inches


ggarrange(RFplot2, MLRplot, common.legend = FALSE)


dev.off()


```



# Total data set model comparison:
```{r}


all.final.model <- lm(TEP ~ Log_Chl + MLD + Temperature + Nitrate, 
                          data = as.data.frame(scale(
                            all.learning.ds[, sapply(all.learning.ds, is.numeric)]))) #ESNP

summary(all.final.model)
car::vif(all.final.model)  # multi-collinearity
ols_test_normality(resid(all.final.model))
plot(all.final.model)


all.final.model2 <- lm(TEP ~ Log_Chl + POC + MLD + DO_Prod + Temperature + Nitrate, 
                          data = as.data.frame(scale(
                            all.learning.ds[, sapply(all.learning.ds, is.numeric)]))) #ESNP

all.final.model3 <- lm(TEP ~ Log_Chl + MLD + DO_Prod + Temperature + Nitrate, 
                          data = as.data.frame(scale(
                            all.learning.ds[, sapply(all.learning.ds, is.numeric)]))) #ESNP


AIC(all.final.model, all.final.model2, all.final.model3)

```



# All data: MLR
```{r}

# First, derive multiple linear regression models (MLRs) from the dataset:

# Use Final model versions 1 & 2 from above chunk:


# List models:
model_list.all <- list(all.final.model, all.final.model2)
model_names.all <- c("all", "all")
model_level.all <- c("all", "all2")


CoefficientPlot <- function(models, alpha = 0.1, 
                            modelnames = model_names.all, modellevel = model_level.all){
  # models must be a list()

  Multiplier <- qnorm(1 - alpha / 2)
  CoefficientTables <- lapply(models, function(x){summary(x)$coef})
  TableRows <- unlist(lapply(CoefficientTables, nrow))

  if(modelnames[1] == ""){
    ModelNameLabels <- rep(paste("Model", 1:length(TableRows)), TableRows)
    } else {
    ModelNameLabels <- rep(modelnames, TableRows)
    }
  
    if(modellevel[1] == ""){
    ModelLevelLabels <- rep(paste("Level", 1:length(TableRows)), TableRows)
    } else {
    ModelLevelLabels <- rep(modellevel, TableRows)
    }

  MatrixofModels <- cbind(do.call(rbind, CoefficientTables), ModelNameLabels, ModelLevelLabels)
  MatrixofModels <- data.frame(cbind(rownames(MatrixofModels), MatrixofModels, ModelLevelLabels))
  colnames(MatrixofModels) <- c("IV", "Estimate", "StandardError", 
                                "TValue", "PValue", "ModelName", "ModelLevel")
  MatrixofModels$ModelLevel = factor(MatrixofModels$ModelLevel, 
                                    levels = c("all", "all2"))
  MatrixofModels$ModelName = factor(MatrixofModels$ModelName, 
                                    levels = c("all", "all2"))
  #MatrixofModels$IV <- factor(MatrixofModels$IV, levels = MatrixofModels$IV)
  #MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})
  MatrixofModels[, -c(1, 6, 7)] <- apply(MatrixofModels[, -c(1, 7)], 2, function(x){as.numeric(as.character(x))})

  OutputPlot <- qplot(IV, Estimate, 
                      ymin = Estimate - Multiplier * StandardError,
                      ymax = Estimate + Multiplier * StandardError,
                      data = MatrixofModels, geom = "pointrange", 
                      ylab = NULL, xlab = NULL, 
                      group = ModelLevel, shape = ModelLevel, alpha = ModelLevel,
                      colour = driver, fill = driver, size = I(1)) +
    geom_hline(yintercept = 0, lwd = I(7/12), colour = I(hsv(0/12, 7/12, 7/12)), 
              alpha = I(5/12)) +
    coord_flip() + theme_bw() +
    theme(panel.grid.major.y = element_blank(), legend.title = element_blank()) +
    ggtitle("Total Scope: Multiple Linear Regression") +
    #geom_point(colour = 'black') +
    scale_x_discrete(limits = c(
                                "Area", "Salinity", "WS", "MLD",
                                "Silicate", "Phosphate", "Nitrate", "Temperature",
                                "DO_Prod", "POC", "Log_Chl"), 
                     labels = c(
                              "Area", "Salinity", "Wind speed", "MLD",
                              "Silicic Acid","Phosphate", "Nitrate", "Temperature",
                              "AOU", "POC", "Chlorophyll a")) +
    scale_shape_manual(values = c("all" = 19, "all2" = 1)) + 
    scale_fill_manual(values = c("Source" = "yellow4", "Direct Environ." = "blue", 
                                 "Indirect Environ." = "darkcyan")) +
    scale_colour_manual(values = c("Source" = "yellow4", "Direct Environ." = "blue", 
                                 "Indirect Environ." = "darkcyan")) +
    xlab("") +
    ylim(-0.3, 0.75) +
    scale_alpha_manual(values = c(1, 1)) +
    theme(legend.title = element_blank(), legend.position = "none", 
          axis.text.y = element_blank(), 
          plot.title = element_text(size = 6, face = "bold")) +
    xlab("")
  
  return(OutputPlot)
}


MLRplot.all <- CoefficientPlot(model_list.all)

# Choose the color scheme corresponding to each parameter:
# driver <- c("Indirect Environ.", "Source", "Indirect Environ.", "Direct Environ.",
#                 "Source", "Source", "Source",
#                 "Source", "Inirect Environ.", "Direct Environ.", "Direct Environ.", "Source")  #  driver categories

driver <- c("Source", "Source", 
            "Indirect Environ.", # MLD
            "Direct Environ.", # Temperature
            "Direct Environ.", # Nitrate
            "Direct Environ.", # Temp
            "Source", # Chl?
            "Source",# POC
            "Indirect Environ.", # MLD
            "Source",  # AOU
            "Direct Environ.", "Direct Environ.") # T, N

# order: Log_Chl, Temperature, Nitrate, DO_Prod, MLD, Log_Chl, Temperature, DO_Prod, MLD, Nitrate, POC, WS,
# Phosphate


all.plots <- ggarrange(RFplot.all, MLRplot.all)




```

# Compare all plots:
```{r}

# print the pdf:
pdf(file = "Coefficients.pdf",   # The directory you want to save the file in
    width = 12, # The width of the plot in inches
    height = 10) # The height of the plot in inches


ggarrange(esnp.plots, all.plots, ncol = 1)


dev.off()




```


